Manuscript

### 2018-10-22 manuscript.Rmd
update plots and Wilcoxon tests w/ new G2P filtering

num_cores <- parallel::detectCores()
options(mc.cores = num_cores, expressions = 5e5)
```

>>>>

# num_cores <- parallel::detectCores()
options(mc.cores = parallel::detectCores(), expressions = 5e5)
```

```{r}
library(bsselectR)
library(data.table)
library(ggpubr)
library(ggsignif)
library(kableExtra)
library(tidyverse)
```

# Functions {.tabset .tabset-fade .tabset-pills}

## Significance thresholding

```{r}
adj_signif <- function(df) {
  # Takes df from compare_means and creates a signif code column for adjusted p-vals
  df$p.signif <- ifelse(df$p.signif == "ns", NA, df$p.signif)
  df$p.signif.adj <- ifelse(df$p.adj <= 0.0001, "****",
                     ifelse(df$p.adj <= 0.001, "***",
                     ifelse(df$p.adj <= 0.01, "**",
                     ifelse(df$p.adj <= 0.05, "*", NA))))
  df$p.short <- formatC(df$p, format = "g", digits =  2)
  df$p.adj.short <- formatC(df$p.adj, format = "g", digits =  2)

  return(df)
}
```

## Wilcoxon tests by drug

```{r}
WilcoxonByDrug <- function(drug, dataset, data_name) {
  use_data <- filter(dataset, Drug == drug)
  sig <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Hugo_Symbol", data = use_data, method = "wilcox.test", p.adjust.method = "BH")
  sig <- adj_signif(sig)
  sig <- sig[order(sig$p),]
  sig$Drug <- drug
  sig$Dataset <- data_name
  return(sig)
}
```

# Data management {.tabset .tabset-fade .tabset-pills}

## D. Charytonowicz cell line converter

This comprehensive cancer cell line information curated by Daniel Charytonowicz.
```{r}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
```

## DepMap cell line metadata

```{r}
ccl_info <- read.delim("./data_munging/DepMap-2018q3-celllines.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_info$Primary.Disease <- gsub("\\\\", "", ccl_info$Primary.Disease)
ccl_info$Primary.Disease <- gsub("Ewings", "Ewing's", ccl_info$Primary.Disease)

# From figshare
crispr_meta <- read.delim("./data_munging/sample_info_18Q3_crispr.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[7] <- "CCLE_Name"
```

## Genomic annotations

### CCLE MAF file

For mutation calling, get paired gene name and cell line fields in a data frame. For this analysis, we don't care about how many mutations there are per gene or what type of mutations there are, so I didn't save more information. I took unique gene-cell line combinations since we only cared about mutation presence/absence. Add a `Mutation_Status` column denoting all entries in MAF files as mutations present in the associated cell lines.
```{r}
maf_raw <- read.delim("./data_munging/CCLE_DepMap_18q3_maf_20180718.txt.gz", header = TRUE, sep = "\t")
# Filter for cell lines in CRISPR screen
maf_raw <- filter(maf_raw, Broad_ID %in% unique(crispr_meta$Broad_ID))
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"

# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "Broad_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))

# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Other", "Mutant")
maf_df <- unique(subset(maf_df, select = c("Hugo_Symbol", "CCLE_Name", "Broad_ID", "Mutation_Status_Nonsilent", "Reference_Allele_Length", "Tumor_Seq_Allele1_Length")))
```

### CCLE copy number file

```{r, eval=FALSE}
cn <- read.delim("./data_munging/public_18Q3_gene_cn_v2.csv.gz", sep = ",", check.names = FALSE, header = TRUE)

# Convert log2 ratios (log2(CN/2)) to CN
cn[2:ncol(cn)] <- lapply(cn[2:ncol(cn)], function(x) 2 * (2 ^ x))

# Remove Entrez gene IDs from colnames
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"

# Melt
cn_melt <- melt(data = cn, id.vars = "Broad_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Hugo_Symbol", value.name = "Copy_Number")

saveRDS(cn_melt, "./data_munging/rds/cn_melt_18Q3.rds", compress = "xz")
```

```{r, eval=FALSE}
cn_melt <- readRDS("./data_munging/rds/cn_melt_18Q3.rds")
```

### CCLE gene expression file

Gene expression data (Reads Per Kilobase of transcript, per Million mapped reads, RPKM).
```{r, eval=FALSE}
ge <- read.delim("./../crispr_lineages_giant_files/CCLE_DepMap_18q3_RNAseq_RPKM_20180718.gct.gz", skip = 2, header = TRUE, sep = "\t", check.names = FALSE)

# Edit columns
ge$Name <- NULL
colnames(ge)[1] <- "Hugo_Symbol"

# Melt
ge_melt <- melt(data = ge, id.vars = "Hugo_Symbol", measure.vars = colnames(ge[2:ncol(ge)]), value.name = "RPKM")
## Split variable column
ge_melt <- with(ge_melt, cbind(Hugo_Symbol, colsplit(variable, pattern = " ", names = c("CCLE_Name", "Broad_ID")), RPKM))
## Remove parentheses around Broad IDs
ge_melt$Broad_ID <- gsub("\\(|\\)", "", ge_melt$Broad_ID)

saveRDS(ge_melt, "./../crispr_lineages_giant_files/ge_melt_18Q3.rds", compress = "xz")
```

```{r, eval=FALSE}
ge_melt <- readRDS("./../crispr_lineages_giant_files/ge_melt_18Q3.rds")
```

## G2P

```{r}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_v2_10_10_2018.tsv", sep = "\t", header = TRUE)
g2p$Drug <- tolower(g2p$Drug)
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p <- subset(g2p, grepl("^CID", g2p$DrugID))

g2p_drug_genes <- as.character(unique(g2p$Drug_Gene[!is.na(g2p$Drug_Gene)]))
g2p_genes <- unique(sapply(strsplit(g2p_drug_genes, "_", fixed = TRUE), function(x) x[2]))
g2p_drugs <- unique(sapply(strsplit(g2p_drug_genes, "_", fixed = TRUE), function(x) x[1]))

g2p_genes
g2p_drugs
# Order and genes taken from pink heatmapo from Daniel's analyses
# g2p_genes_filt <- c("KRAS", "EGFR", "MET", "ALK", "BRAF", "ERBB2", "ABL1", "ROS1", "KIT", "DPYD", "PDGFRA", "RET", "BRCA1", "UGT1A1", "IDH2", "ESR1", "TPMT", "BRCA2", "PDGFRB", "G6PD", "PML", "TSC1", "FLT3", "ERBB3", "CYP19A1", "AKT1", "PTCH1", "MGMT", "DNMT3A", "MYD88", "PDGFB", "TSC2", "NPM1", "JAK2")
```

## CRISPR

### DepMap dependency probabilities

```{r, eval=FALSE}
dep <- read.delim("./data_munging/gene_dependency_18Q3.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(dep) <- gsub(" .*", "", colnames(dep))
```

### Merge CRISPR data and annotations

The latest CRISPR CERES score data (18Q3, August 2018) was pulled from the [DepMap Data Portal](https://depmap.org/portal/download/) (Broad Institute Cancer Dependency Map 2018, Meyers et al. 2017).
```{r, eval=TRUE}
crispr <- read.delim("./data_munging/gene_effect_18Q3.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr) <- gsub(" .*", "", colnames(crispr))
```

Merge annotation data:
```{r, eval=FALSE}
# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr, id.vars = "Broad_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Hugo_Symbol", value.name = "Score")
# Melt dependency probabilities dataset for merging
dep_melt <- melt(dep, id.vars = "Broad_ID", measure.vars = colnames(dep)[2:ncol(dep)], variable.name = "Hugo_Symbol", value.name = "Dep_Prob")
# Merge dependency probabilities
crispr_melt <- merge(crispr_melt, dep_melt, by = c("Broad_ID", "Hugo_Symbol"), all.x = TRUE)
# Merge cell line metadata
crispr_melt <- merge(crispr_melt, ccl_info, by = "Broad_ID", all.x = TRUE)
crispr_melt <- merge(crispr_melt, crispr_meta, by = c("CCLE_Name", "Broad_ID"), all.x = TRUE)
# Merge mutation annotations
crispr_muts <- merge(crispr_melt, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
crispr_muts$Hugo_Symbol <- factor(crispr_muts$Hugo_Symbol)
crispr_muts <- crispr_muts %>% mutate(Mutation_Status_Nonsilent = if_else(is.na(Mutation_Status_Nonsilent), "Other", Mutation_Status_Nonsilent))
# Summarize number of mutant and Other cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Hugo_Symbol) %>%
  summarize(N_Nonsilent_Other = sum(Mutation_Status_Nonsilent == "Other"),
            N_Nonsilent_Mutant = sum(Mutation_Status_Nonsilent == "Mutant"))
# Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Hugo_Symbol")
# Add Color columns
crispr_data$Color_Nonsilent <- ifelse(crispr_data$Mutation_Status_Nonsilent == "Other", "thistle", "darkorchid")
crispr_data$Color_Nonsilent <- factor(crispr_data$Color_Nonsilent)
# Cell line lineages
crispr_data <- merge(crispr_data, ccl_converter, by = c("CCLE_Name", "Broad_ID"), all.x = TRUE)
levels(crispr_data$lineage_name) <- sort(levels(crispr_data$lineage_name), decreasing = TRUE)
# Copy number
crispr_data <- merge(crispr_data, cn_melt, by = c("Hugo_Symbol", "Broad_ID"), all.x = TRUE)
# Gene expression (RPKM)
ge_filt <- filter(ge_melt, Hugo_Symbol %in% unique(crispr_data$Hugo_Symbol) & Broad_ID %in% unique(crispr_data$Broad_ID))
crispr_data <- merge(crispr_data, ge_filt, by = c("Hugo_Symbol", "Broad_ID", "CCLE_Name"), all.x = TRUE)
crispr_data$RPKM_log2 <- log2(crispr_data$RPKM + 0.0001)
saveRDS(crispr_data, "./../crispr_lineages_giant_files/crispr_data_18Q3.rds", compress = "xz")
```

```{r}
crispr_data <- readRDS("./../crispr_lineages_giant_files/crispr_data_18Q3.rds")

crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)

# Filter CRISPR for G2P genes in G2P
crispr_data_g2p <- filter(crispr_data_ptmuts, Hugo_Symbol %in% g2p_genes)
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
```

## Drug screens

```{r, eval=FALSE}
# Drug metadata
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)
maf_df_g2p <- filter(maf_df, Hugo_Symbol %in% g2p_genes)

# CCLE
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "AUC")
ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ccle_data <- merge(ccle_grid, ccle_melt, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, ccl_converter, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
ccle_data$Mutation_Status_Nonsilent <- ifelse(is.na(ccle_data$Mutation_Status_Nonsilent), "Other", ccle_data$Mutation_Status_Nonsilent)
ccle_data$Hugo_Symbol <- factor(ccle_data$Hugo_Symbol)
ccle_data$AUC <- as.numeric(ccle_data$AUC)
ccle_data$Color_Nonsilent <- ifelse(ccle_data$Mutation_Status_Nonsilent == "Other", "palegreen3", "springgreen4")
ccle_data$Color_Nonsilent <- factor(ccle_data$Color_Nonsilent)
ccle_data <- filter(ccle_data, !is.na(ccle_data$AUC))
ccle_data$Drug_Gene <- paste0(ccle_data$Drug, "_", ccle_data$Hugo_Symbol)
saveRDS(ccle_data, "./data_munging/rds/ccle_data_18Q3.rds", compress = "xz")

# CTRP
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.vars = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC")
ctrp_grid <- expand.grid("accession_id" = unique(ctrp_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ctrp_data <- merge(ctrp_grid, ctrp_melt, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, ccl_converter, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
ctrp_data$Mutation_Status_Nonsilent <- ifelse(is.na(ctrp_data$Mutation_Status_Nonsilent), "Other", ctrp_data$Mutation_Status_Nonsilent)
ctrp_data$Hugo_Symbol <- factor(ctrp_data$Hugo_Symbol)
ctrp_data$AUC <- as.numeric(ctrp_data$AUC)
ctrp_data$Color_Nonsilent <- ifelse(ctrp_data$Mutation_Status_Nonsilent == "Other", "slategray3", "steelblue4")
ctrp_data$Color_Nonsilent <- factor(ctrp_data$Color_Nonsilent)
ctrp_data <- filter(ctrp_data, !is.na(ctrp_data$AUC))
ctrp_data$Drug_Gene <- paste0(ctrp_data$Drug, "_", ctrp_data$Hugo_Symbol)
saveRDS(ctrp_data, "./data_munging/rds/ctrp_data_18Q3.rds", compress = "xz")

# GDSC
gdsc <- read.delim("./data_munging/data_drug_auc_gdsc_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
gdsc <- merge(drug_meta, gdsc, by = "CID")
gdsc_melt <- melt(gdsc, id.vars = c("CID", "Drug"), measure.vars = colnames(gdsc)[3:ncol(gdsc)], variable.name = "accession_id", value.name = "AUC")
gdsc_grid <- expand.grid("accession_id" = unique(gdsc_melt$accession_id), "Hugo_Symbol" = g2p_genes)
gdsc_data <- merge(gdsc_grid, gdsc_melt, by = "accession_id", all.x = TRUE)
gdsc_data <- merge(gdsc_data, ccl_converter, by = "accession_id", all.x = TRUE)
gdsc_data <- merge(gdsc_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
gdsc_data$Mutation_Status_Nonsilent <- ifelse(is.na(gdsc_data$Mutation_Status_Nonsilent), "Other", gdsc_data$Mutation_Status_Nonsilent)
gdsc_data$Hugo_Symbol <- factor(gdsc_data$Hugo_Symbol)
gdsc_data$AUC <- as.numeric(gdsc_data$AUC)
gdsc_data$Color_Nonsilent <- ifelse(gdsc_data$Mutation_Status_Nonsilent == "Other", "paleturquoise3", "turquoise4")
gdsc_data$Color_Nonsilent <- factor(gdsc_data$Color_Nonsilent)
gdsc_data <- filter(gdsc_data, !is.na(gdsc_data$AUC))
gdsc_data$Drug_Gene <- paste0(gdsc_data$Drug, "_", gdsc_data$Hugo_Symbol)
saveRDS(gdsc_data, "./data_munging/rds/gdsc_data_18Q3.rds", compress = "xz")
```

```{r}
ccle_data <- readRDS("./data_munging/rds/ccle_data_18Q3.rds")
ccle_data_ptmuts <- filter(ccle_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ccle_data_ptmuts$Keep_PtMut_Tests <- ifelse(ccle_data_ptmuts$Reference_Allele_Length == 1 | is.na(ccle_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ccle_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ccle_data_ptmuts <- filter(ccle_data_ptmuts, Keep_PtMut_Tests == TRUE)

ctrp_data <- readRDS("./data_munging/rds/ctrp_data_18Q3.rds")
ctrp_data_ptmuts <- filter(ctrp_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ctrp_data_ptmuts$Keep_PtMut_Tests <- ifelse(ctrp_data_ptmuts$Reference_Allele_Length == 1 | is.na(ctrp_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ctrp_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ctrp_data_ptmuts <- filter(ctrp_data_ptmuts, Keep_PtMut_Tests == TRUE)

gdsc_data <- readRDS("./data_munging/rds/gdsc_data_18Q3.rds")
gdsc_data_ptmuts <- filter(gdsc_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
gdsc_data_ptmuts$Keep_PtMut_Tests <- ifelse(gdsc_data_ptmuts$Reference_Allele_Length == 1 | is.na(gdsc_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(gdsc_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
gdsc_data_ptmuts <- filter(gdsc_data_ptmuts, Keep_PtMut_Tests == TRUE)

# Filter drug screen datasets for G2P drug-gene associations
ccle_data_g2p <- filter(ccle_data_ptmuts, Drug_Gene %in% g2p_drug_genes)
ctrp_data_g2p <- filter(ctrp_data_ptmuts, Drug_Gene %in% g2p_drug_genes)
gdsc_data_g2p <- filter(gdsc_data_ptmuts, Drug_Gene %in% g2p_drug_genes)
```

# Gene mutation frequency

```{r}
crispr_data_g2p_summ <- unique(crispr_data_g2p[, c("Hugo_Symbol", "N_Nonsilent_Mutant", "N_Nonsilent_Other")])
crispr_data_g2p_summ$Fraction_Mutant <- crispr_data_g2p_summ$N_Nonsilent_Mutant / (crispr_data_g2p_summ$N_Nonsilent_Mutant + crispr_data_g2p_summ$N_Nonsilent_Other)
crispr_data_g2p_summ$Percent_Mutant <- crispr_data_g2p_summ$Fraction_Mutant * 100
# write.table(crispr_data_g2p_summ, file = "~/Desktop/crispr_data_g2p_summ_20181003.tsv", quote = FALSE, sep = "\t")
```

# Wilcoxon tests

## CRISPR: Gene

```{r, eval=FALSE}
crispr_signif_g2p_gene <- compare_means(Score ~ Mutation_Status_Nonsilent, group.by = c("Hugo_Symbol"), data = crispr_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_g2p_gene <- adj_signif(crispr_signif_g2p_gene)
crispr_signif_g2p_gene <- crispr_signif_g2p_gene[order(crispr_signif_g2p_gene$p),]
saveRDS(crispr_signif_g2p_gene, "./data_munging/rds/crispr_signif_g2p_gene.rds")
```

```{r, eval=TRUE}
crispr_signif_g2p_gene <- readRDS("./data_munging/rds/crispr_signif_g2p_gene.rds")

crispr_signif_g2p_20gene <- as.character(crispr_signif_g2p_gene$Hugo_Symbol[0:20])

knitr::kable(filter(crispr_signif_g2p_gene, p < 0.1)[, c("Hugo_Symbol", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results for Level A point mutations, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=7, fig.width=12}
crispr_data_signif_g2p_gene_filt <- filter(crispr_data_ptmuts, Hugo_Symbol %in% crispr_signif_g2p_20gene)
crispr_data_signif_g2p_gene <- merge(crispr_data_signif_g2p_gene_filt, crispr_signif_g2p_gene, by = "Hugo_Symbol")
crispr_data_signif_g2p_gene$Hugo_Symbol <- ordered(crispr_data_signif_g2p_gene$Hugo_Symbol, levels = crispr_signif_g2p_20gene)
crispr_data_signif_g2p_gene_color <- as.character(crispr_data_signif_g2p_gene$Color_Nonsilent)
names(crispr_data_signif_g2p_gene_color) <- crispr_data_signif_g2p_gene$Mutation_Status_Nonsilent
crispr_data_signif_g2p_gene <- merge(crispr_data_signif_g2p_gene, crispr_data_signif_g2p_gene %>% group_by(Hugo_Symbol) %>% summarize(Text_y = max(Score)), by = "Hugo_Symbol")

crispr_data_signif_g2p_gene_text <- unique(data.frame(p.signif = crispr_data_signif_g2p_gene$p.signif, p.signif.adj = crispr_data_signif_g2p_gene$p.signif.adj,  Hugo_Symbol = crispr_data_signif_g2p_gene$Hugo_Symbol, Text_y = crispr_data_signif_g2p_gene$Text_y))

crispr_data_signif_g2p_gene_plot <- ggplot(data = crispr_data_signif_g2p_gene, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, nrow = 2, scales = "free_y") +
  geom_jitter(aes(color = Mutation_Status_Nonsilent), alpha = 0.5, width = 0.3) +
  geom_boxplot(color = "black", fill = NA, outlier.shape = NA) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = crispr_data_signif_g2p_gene_color) + 
  geom_text(data = crispr_data_signif_g2p_gene_text, mapping = aes(x = 1.5, y = Text_y, label = p.signif), nudge_y = 0.2) +
  theme(legend.position = "top") +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES scores for select Level A G2P genes",
       subtitle = "Genes are sorted by increasing p-value.")
crispr_data_signif_g2p_gene_plot
# Genes with adjusted p-values < 0.3 are shown.\nBH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\n
# ggsave("./plots_18Q3/manuscript/crispr_data_signif_g2p_gene_plot.png", crispr_data_signif_g2p_gene_plot, width = 12, height = 7, units = "in")
```

## CRISPR: Gene and lineage

```{r, eval=FALSE}
crispr_signif_g2p_lineage <- compare_means(Score ~ Mutation_Status_Nonsilent, group.by = c("Hugo_Symbol", "group_general_lineage_name"), data = crispr_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_g2p_lineage <- adj_signif(crispr_signif_g2p_lineage)
crispr_signif_g2p_lineage <- crispr_signif_g2p_lineage[order(crispr_signif_g2p_lineage$p),]
saveRDS(crispr_signif_g2p_lineage, "./data_munging/rds/crispr_signif_g2p_lineage.rds")

# write.table(crispr_signif_g2p_lineage, file = "~/Desktop/crispr_signif_g2p_lineage.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=TRUE}
crispr_signif_g2p_lineage <- readRDS("./data_munging/rds/crispr_signif_g2p_lineage.rds")

knitr::kable(filter(crispr_signif_g2p_lineage, p < 0.01)[, c("Hugo_Symbol", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing non-silent mutant vs other cell lines by lineage, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=15, fig.width=15, eval=FALSE}
crispr_data_signif_g2p_lineage <- merge(crispr_data_signif_g2p_gene_filt, crispr_signif_g2p_lineage, by = c("Hugo_Symbol", "group_general_lineage_name"))
crispr_data_signif_g2p_lineage$Hugo_Symbol <- ordered(crispr_data_signif_g2p_lineage$Hugo_Symbol, levels = crispr_signif_g2p_20gene)
crispr_data_signif_g2p_lineage_color <- as.character(crispr_data_signif_g2p_lineage$Color_Nonsilent)
names(crispr_data_signif_g2p_lineage_color) <- crispr_data_signif_g2p_lineage$Mutation_Status_Nonsilent
crispr_data_signif_g2p_lineage <- merge(crispr_data_signif_g2p_lineage, crispr_data_signif_g2p_lineage %>% group_by(Hugo_Symbol) %>% summarize(Text_y = max(Score)), by = "Hugo_Symbol")

crispr_data_signif_g2p_lineage_text <- unique(data.frame(p.signif = crispr_data_signif_g2p_lineage$p.signif, p.signif.adj = crispr_data_signif_g2p_lineage$p.signif.adj,  group_general_lineage_name = crispr_data_signif_g2p_lineage$group_general_lineage_name, Hugo_Symbol = crispr_data_signif_g2p_lineage$Hugo_Symbol, Text_y = crispr_data_signif_g2p_lineage$Text_y))

crispr_data_signif_g2p_lineage_plot <- ggplot(data = crispr_data_signif_g2p_lineage, aes(x = group_general_lineage_name, y = Score)) +
  facet_wrap(~ Hugo_Symbol, nrow = 4, scales = "free_y") +
  geom_point(aes(color = Mutation_Status_Nonsilent), alpha = 0.5) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = crispr_data_signif_g2p_lineage_color) +
  geom_text(data = crispr_data_signif_g2p_lineage_text, mapping = aes(x = group_general_lineage_name, y = Text_y, label = p.signif), angle = 90, nudge_y = 0.1) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1, size = 10)) +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES score by cell line lineage for significant Level A G2P genes",
       subtitle = "Genes with adjusted p-values < 0.3 are shown.\nBH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\nGenes are sorted by increasing lineage-agnostic p-value.")
crispr_data_signif_g2p_lineage_plot

# ggsave("./plots_18Q3/manuscript/crispr_crispr_data_signif_g2p_lineage_plot.png", crispr_data_signif_g2p_lineage_plot, width = 12, height = 15, units = "in")
```

```{r}
crispr_data_signif_g2p_lineage_summ <- crispr_data_signif_g2p_lineage %>% group_by(Hugo_Symbol, group_general_lineage_name, Mutation_Status_Nonsilent) %>% tally()
# write.table(crispr_data_signif_g2p_lineage_summ, file = "~/Desktop/crispr_data_signif_g2p_lineage_summ_20181005.tsv", quote = FALSE, sep = "\t")

knitr::kable(crispr_data_signif_g2p_lineage_summ, caption = "Mutant/Other counts for gene-lineage combinations") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## CCLE: G2P gene and drug

```{r, eval=FALSE}
ccle_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = ccle_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
ccle_signif_g2p <- adj_signif(ccle_signif_g2p)
ccle_signif_g2p <- ccle_signif_g2p[order(ccle_signif_g2p$p),]
saveRDS(ccle_signif_g2p, "./data_munging/rds/ccle_signif_g2p_gene.rds")
```

```{r, eval=TRUE}
ccle_signif_g2p <- readRDS("./data_munging/rds/ccle_signif_g2p_gene.rds")

knitr::kable(ccle_signif_g2p[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: Wilcoxon test results for Level A point mutations, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## CTRP: G2P gene and drug

```{r, eval=FALSE}
ctrp_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = ctrp_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
ctrp_signif_g2p <- adj_signif(ctrp_signif_g2p)
ctrp_signif_g2p <- ctrp_signif_g2p[order(ctrp_signif_g2p$p),]
saveRDS(ctrp_signif_g2p, "./data_munging/rds/ctrp_signif_g2p_gene.rds")
```

```{r, eval=TRUE}
ctrp_signif_g2p <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")

knitr::kable(ctrp_signif_g2p[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CTRP: Wilcoxon test results for Level A point mutations, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## GDSC: G2P gene and drug

```{r, eval=FALSE}
gdsc_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = gdsc_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
gdsc_signif_g2p <- adj_signif(gdsc_signif_g2p)
gdsc_signif_g2p <- gdsc_signif_g2p[order(gdsc_signif_g2p$p),]
saveRDS(gdsc_signif_g2p, "./data_munging/rds/gdsc_signif_g2p_gene.rds")
```

```{r, eval=TRUE}
gdsc_signif_g2p <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")

knitr::kable(gdsc_signif_g2p[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "GDSC: Wilcoxon test results for Level A point mutations, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

# Boxplots

Mutation status boxplots for genes identified in G2P.

## CRISPR

```{r fig.height=5, fig.width=12, eval=FALSE}
crispr_data_g2p_color <- as.character(crispr_data_g2p$Color_Nonsilent)
names(crispr_data_g2p_color) <- crispr_data_g2p$Mutation_Status_Nonsilent

crispr_label_text <- data.frame(p.signif.adj = crispr_signif_g2p_gene$p.signif.adj, p.signif = crispr_signif_g2p_gene$p.signif, Hugo_Symbol = crispr_signif_g2p_gene$Hugo_Symbol)
crispr_label_text <- filter(crispr_label_text, Hugo_Symbol %in% g2p_genes)
crispr_label_text$Hugo_Symbol <- factor(crispr_label_text$Hugo_Symbol, levels = g2p_genes)
